---
title: "Netflix Shows"
author:
  - Anthony Lipphardt
date: 2021-04-26
output: github_document

knit: (function(input, ...) { rmarkdown::render(input, output_file = 'README.md') })

---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) event!

Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.

While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library("tidytuesdayR")

library("tidyverse")
library("lubridate")
library("scales")
library("ggthemes")


```

# Loading the Weekly Dataset

Download the weekly data and make available in the `netflix` object.

[Netflix Shows](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-20/readme.md)

```{r Load}

# Loading Data for the First Time
# tuesdata <- tidytuesdayR::tt_load('2021-04-20')
# netflix <- tuesdata$netflix
# write_csv(netflix, 'netflix.csv')

netflix <- read_csv('netflix.csv')

```


# Sample Data

Take an initial look at the format of the data available.

```{r Glimpse}

head(netflix, 10)

```

# Wrangling

Explore the data and process it into prepared dataframes for visualization.

```{r Wrangle}

netflix_shows_added_by_genre = netflix %>%
  
  # Keep only TV shows
  select(type, listed_in, date_added) %>%
  filter(type == "TV Show") %>%
  
  # Create an entry for each TV Show/Genre pairing
  mutate(listed_in = str_split(listed_in, ",")) %>%
  unnest(listed_in) %>%
  mutate(listed_in = str_trim(listed_in)) %>%
  
  # Retain TV Show specific categories
  filter(str_starts(listed_in, "TV"), listed_in != 'TV Shows') %>%
  mutate(listed_in = str_replace(listed_in, "TV ", "")) %>%
  
  # Parse year from date_added and remove NA dates
  mutate(year = year(parse_date_time(date_added, "b d, y"))) %>% 
  filter(!is.na(year)) %>%
  
  # Get number of of shows added for each genre-year pair
  group_by(listed_in, year) %>%
  summarise(n = n()) %>%
  
  # Convert year to POSIXct date for axis formatting
  mutate(year = parse_date_time(year, "y"))
  


```


# Visualization(s)

Using your processed dataset, create your unique visualization(s).

```{r Visualize, fig.width=16, fig.height=9}

  # Plot number of shows added per year, by genre
  plot1 = netflix_shows_added_by_genre %>%
  ggplot(aes(x=year)) +
  geom_line(aes(y=n, color = listed_in), size=1.2) +
  labs(title = "TV Shows Added to   Netflix   Over Time, by Genre",
       x = "Year Show Was Added",
       y = "Number of Shows Added",
       color = "Genre") + 
  
  # Additional formatting/themes
  scale_x_datetime(labels = date_format("%Y")) +
  theme_few() + 
  scale_color_tableau() +
  theme(plot.title = element_text(size=20, face="bold"))

  plot1
  
```

# Saving Image(s)

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(plot = plot1,
  filename = "netflix-shows-added-by-genre.png",
  width=16,
  height=9,
  device = "png")

```
