---
title: "CEO Departures"
author: Developed by Anthony Lipphardt
output: github_document

knit: (function(input, ...) { rmarkdown::render(input, output_file = 'README.md') })

---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) event!

Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.

While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library("tidytuesdayR")

library("tidyverse")
library("stringi")
library("lubridate")
library("scales")
library("ggthemes")


```

# Loading the Weekly Dataset

Download the weekly data and make available in the `departures` object.

[CEO Departures](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-27/readme.md)

```{r Load}

# Loading Data for the First Time
# tuesdata <- tidytuesdayR::tt_load('2021-04-27')
# write_csv(tuesdata$departures, "departures.csv")

departures <- read_csv('departures.csv')

```


# Sample Data

Take an initial look at the format of the data available.

```{r Glimpse}

glimpse(head(departures, 10))

```

# Wrangling

Explore the data and process it into prepared dataframes for visualization.

```{r Wrangle}

male_pronouns = c("He", "he", "Him", "him", "His", "his", "Mr", "Mister")
female_pronouns = c("She", "she", "Her", "her", "Hers", "hers", "Miss", "Mrs", "Ms")

pronoun_regex = function(pronouns){
  results = stri_join_list(lapply(list(pronouns), 
                        FUN = function(pronoun){ 
                          return(paste0("\\b", pronoun, "\\b")) 
                          }), sep = "|")
  return(results)
}


departures %>%
  mutate(gender = case_when(
    str_detect(notes, regex(pronoun_regex(male_pronouns))) ~ 'M',
    str_detect(notes, regex(pronoun_regex(female_pronouns))) ~ 'F',
    TRUE ~ 'NA'
  )) %>%
  filter(gender != 'NA', !is.na(departure_code)) %>%
  group_by(gender) %>%
  summarize(count = n()) 
   

departure_by_gender = departures %>%
  mutate(gender = case_when(
    str_detect(notes, regex(pronoun_regex(male_pronouns))) ~ 'M',
    str_detect(notes, regex(pronoun_regex(female_pronouns))) ~ 'F',
    TRUE ~ 'NA'
  )) %>%
  filter(gender != 'NA', !is.na(departure_code)) %>%
  group_by(gender, departure_code) %>%
  summarize(count = n()) %>%
  mutate(gender_percent = count/sum(count)*100) %>%
  mutate(gender_percent = ifelse(gender == "M", gender_percent,-1*gender_percent)) %>%
  filter(departure_code < 8) %>%
  mutate(departure_code = case_when(
    departure_code == 1 ~ "Death",
    departure_code == 2 ~ "Illness",
    departure_code == 3 ~ "Dismissed, Job Performance",
    departure_code == 4 ~ "Dismissed, Legal Violations",
    departure_code == 5 ~ "Retired",
    departure_code == 6 ~ "New Opportunity",
    departure_code == 7 ~ "Other",
    departure_code == 8 ~ "Missing",
    departure_code == 9 ~ "Error"
  ))



```


# Visualization(s)

Using your processed dataset, create your unique visualization(s).

```{r Visualize, fig.width=16, fig.height=9}

plot1 = departure_by_gender %>%
  arrange(desc(gender_percent)) %>%
  mutate(departure_code = factor(departure_code, levels=departure_code)) %>%
  ggplot(aes(x=departure_code, y=gender_percent, fill=gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Reasons for CEO Departure, by Gender",
       subtitle = "Based on 128 Female CEOs and 4964 Male CEOs",
       caption = "Source: DataIsPlural, 2021",
       x = "",
       y = "Percentage of CEO Departures",
       fill = "Gender") +   
  ylim(-60,60) +
  coord_flip() +
  theme_few() + 
  scale_color_few() +  
  theme(plot.title = element_text(size=20, face="bold"),
        plot.subtitle = element_text(size=18),
        plot.caption= element_text(size=14),
        axis.text.x = element_text(size=16),
        axis.title.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title.y = element_text(size=16),
        )

plot1
  
```

# Saving Image(s)

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(plot = plot1,
  filename = "reasons-for-ceo-departure-by-gender.png",
  width=16,
  height=9,
  device = "png")


```
