---
title: "Animal Rescues"
author: Developed by Anthony Lipphardt
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = 'README.md') })
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) event!

Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.

While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.

The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!

As such we encourage everyone of all skills to participate!

```{r setup, warning=FALSE, results = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidytuesdayR)
library(tidyverse)

library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

library(scales)
library(glue)
library(ggtext)
library(Cairo)
library(extrafont)

library(patchwork)

library(lubridate)

library(tidytext)
library(igraph)    
library(ggraph)      

library(grid)
library(png)

loadfonts(device = "win", quiet = TRUE)

```

# Loading the Weekly Dataset

Download the weekly data and make it available in the `animal_rescues` object.

[Animal Rescues](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-06-29/readme.md)

```{r Load}

# Loading Data for the First Time
# tuesdata <- tidytuesdayR::tt_load(2021, week = 27)
# animal_rescues <- tuesdata$animal_rescues
# write_csv(animal_rescues, 'animal_rescues.csv')

animal_rescues <- read_csv('animal_rescues.csv') 

```


# Sample Data

Take an initial look at the format of the data available.

```{r Glimpse}

head(glimpse(animal_rescues), 10)


```

# Wrangling and Exploration

Checking unique values for categoricals. 
* **incident_type** shows the dataset only contains incidents from the London Fire Brigade categorized as 'Special Services' which are animal rescues or assistance. 
* **final_description** contains the detailed description of the reason for the distress call, which could be interesting for text mining.
* **animal_group_parent** includes the type of animal involved for the distress call. It could be interesting to look at the types of rescues for different types of animals (dog vs cat vs bird)
* **originof_call** lists whether the distress call was placed by a person or emergency responder.
* **property_type** contains detailed information on the property where the incident occurred
* **property_category** is a broader category that the property_type falls into
* **special_service_type_category** lists whether the service was assistance or rescue and whether the rescue occurred from height, below ground, or water.
* **special_service_type** provides more detailed information on the service type such as whether the animal was domesticated or wild.
* **ward** and **borough** are location specific information and has very high cardinality.


```{r categorical_uniques}

unique_categoricals = function(column){
  column %>% as_tibble() %>% count(value, sort = TRUE)
}

animal_rescues %>%
  select(type_of_incident,
         final_description,
         animal_group_parent,
         originof_call,
         property_type,
         property_category,
         special_service_type_category,
         special_service_type,
         ward,
         borough,
         ) %>%
  map(unique_categoricals)


```

Given that we have location information, it might also be interesting to see the level of coverage on a map. We'll start with a hexbin map to see where calls are concentrated. Of the 7544 incidents, there are 3699 locations that appear to have valid latitude and longitude data. We can see from the hexbin map there is fairly good coverage within London. So this may be a good candidate for visualization.

```{r hexbin_map, fig.width = 9, fig.height = 5}

rescue_locations = animal_rescues %>%
  select(latitude, longitude) %>%
  filter(latitude != "NULL" & longitude != "NULL") %>%
  mutate(latitude = as.numeric(latitude),
         longitude = as.numeric(longitude)) %>% 
  filter(latitude > 0)


world <- ne_countries(scale = "medium", returnclass = "sf")


uk_map = ggplot(data = world %>% filter(name %in% c("United Kingdom", "Ireland"))) +
  geom_sf(fill = "#EEFCEB") +
  coord_sf(xlim = c(-12,3), ylim = c(49, 59), expand = FALSE) + 
  theme_minimal() + 
  geom_hex(data = rescue_locations, aes(x=longitude, y=latitude), bins = 40, show.legend = FALSE) +
  labs(title = "United Kingdom and Ireland",
       x = "Longitude",
       y = "Latitude") + 
  
  annotate(
    geom = "text",
    x = -2,
    y = 52,
    hjust = 0,
    label = "London",
    color = "#333333",
    size = 4,
    family = "Bahnschrift"
  ) + 
  
  annotate(
    geom = "curve",
    size = 0.7,
    x = -1.3,
    xend = -0.6,    
    y = 51.85,
    yend = 51.5,
    colour = "black",
    arrow = arrow(angle = 30, length = unit(0.02,"npc"), type = "closed", ends = "last"),
    lty = 1
  ) +
  
  theme(text = element_text(family = "Bahnschrift"),
        plot.title = element_text(size = 18))

zoom_in = ggplot(data = world %>% filter(name %in% c("United Kingdom", "Ireland"))) +
  geom_sf(fill = "#EEFCEB") +
  coord_sf(xlim = c(-0.6, 0.5), ylim = c(51.2, 51.7), expand = FALSE) + 
  theme_minimal() + 
  geom_hex(data = rescue_locations, aes(x=longitude, y=latitude), bins = 40, show.legend = TRUE) +
  labs(title = "London",
       x = "Longitude",
       y = "Latitude",
       fill = "Incidents") +
  
  theme(text = element_text(family = "Bahnschrift"),
        plot.title = element_text(size = 18))


uk_map + zoom_in

```

Looking at incidents over time. There definitely appears to be a seasonality in the number of incidents reports.

```{r incidents_over_time, fig.height = 6, fig.width = 9}

animal_rescues %>%
  mutate(date_time_of_call=dmy_hm(date_time_of_call)) %>%
  select(date_time_of_call, animal_group_parent) %>%
  group_by(year = year(date_time_of_call), month = month(date_time_of_call), animal_group_parent) %>%
  count() %>%
  mutate(date = as.Date(glue("{month}/1/{year}"), format = "%m/%d/%Y")) %>%
  filter(animal_group_parent %in% c("Cat", "Bird", "Dog", "Fox", "Horse")) %>%
  ggplot(aes(x=date, y=n)) +
  geom_area(aes(fill = animal_group_parent), size = 1.25, alpha = 0.7) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") + 
  labs(title = "Incidents for Animal Groups Over Time",
       x = "",
       y = "Number of Incidents",
       fill = "") +
  theme(text = element_text(family = "Bahnschrift"),
        plot.title = element_text(size = 18))

```

Let's take a closer look at the **final_description** field. It might be useful to look at most frequent bigrams to determine common word pairs in this field. Network graphs are a useful visualization to show relationships in ngrams.

```{r bigrams, fig.width = 10, fig.height = 6}


visualize_bigrams = function(tbl, animal, color){

  set.seed(2021)
  
  animal_bigrams = animal_rescues %>%
    filter(animal_group_parent == animal,
           final_description != "NULL") %>%
    select(final_description) %>%
    unnest_tokens(bigram, final_description, token = "ngrams", format="text", n=2) %>%
    filter(!is.na(bigram)) 
  
  bigram_graph = animal_bigrams %>%
    count(bigram, sort = TRUE) %>%
    select(bigram, n) %>%
    separate(bigram, c("word1","word2"), sep = " ") %>%
    distinct() %>%
    filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word,
           !is.na(word1)) %>%
    slice_max(n = 25, n) %>%
    arrange(desc(n)) %>%
    mutate(n = log10(n)) %>%
    graph_from_data_frame()
  
  
  ggraph(bigram_graph, layout = "fr") +
      theme_void() +
      geom_node_point(color = color, 
                    size = 4) +
      geom_edge_link(aes(edge_alpha = n), 
                   show.legend = FALSE,
                   arrow = grid::arrow(type = "closed", length = unit(.15, "inches")), 
                   end_cap = circle(.07, 'inches')) +
      geom_node_text(aes(label = name), 
                   vjust = 1, 
                   hjust = 1,
                   size = 3) + 
    theme(plot.background = element_rect(fill = color, color = color),
          plot.margin = margin(t = 10, l = 10, r = 10, b = 10),
          panel.background = element_rect(fill = "white", color = color))
    
}



cat = animal_rescues %>% visualize_bigrams("Cat", "#63615B")
bird = animal_rescues %>% visualize_bigrams("Bird", "#3F6CB2")
dog = animal_rescues %>% visualize_bigrams("Dog", "#DBCC83")

cat / bird / dog


```

# Visualization

Let's put it all together into a cohesive visualization.

```{r visualize}

layout <- c(
        area(t = 1,
             b = 5,
             l = 1,
             r = 7),
        area(t = 1,
             b = 5,
             l = 8,
             r = 18))


#################################
#          CAT PLOT             #
#################################

cat_png = readPNG("cat.png")
cat_color = "#63615B"

cat_text = 
ggplot() + 
  geom_textbox(aes(x =  30, y = 45, label = "<span style='font-size: 36pt;'><strong>Cats</strong></span><br/><span style='font-size: 24pt;'>London Fire Brigade responded to <strong>3,649 incidents</strong> between 2009 and 2021. Cats are often trapped in areas of the house such as under the floor boards, in the washing machine, on window ledges, or in cars.</span>"), fill = cat_color, box.colour = cat_color, width = unit(400, "pt"), hjust = 0, vjust = 1, family = "Bahnschrift") +
  scale_x_continuous(limits = c(0,100)) +
  scale_y_continuous(limits = c(0,50)) +
  theme_void() +
  theme(plot.background = element_rect(fill = cat_color, color = cat_color),
        panel.background = element_rect(fill = cat_color, color = cat_color)) +
  
  annotation_custom(grob = rasterGrob(cat_png, 
                                     interpolate=TRUE,
                                     hjust = 0.5,
                                     vjust = 0.5), 
                   xmin = 0, xmax = 25, ymin = 0, ymax = 50) 


cat_plot = cat + cat_text  + plot_layout(design = layout)


#################################
#          BIRD PLOT            #
#################################

bird_png = readPNG("bird.png")
bird_color = "#3F6CB2"

bird_text = ggplot() + 
  geom_textbox(aes(x =  30, y = 45, label = "<span style='font-size: 36pt;'><strong>Birds</strong></span><br/><span style='font-size: 24pt;'>London Fire Brigade responded to <strong>1,530 incidents</strong> between 2009 and 2021. Birds are often trapped when their legs become stuck in netting or wire. Common locations include underneath railway bridges or on rooftops. Incidents include both domesticated and wild birds.</span>"), fill = bird_color, box.colour = bird_color, width = unit(400, "pt"), hjust = 0, vjust = 1, family = "Bahnschrift", color  = "white") +
  scale_x_continuous(limits = c(0,100)) +
  scale_y_continuous(limits = c(0,50)) +
  theme_void() +
  theme(plot.background = element_rect(fill = bird_color, color = bird_color),
        panel.background = element_rect(fill = bird_color, color = bird_color)) +
  
  annotation_custom(grob = rasterGrob(bird_png, 
                                     interpolate=TRUE,
                                     hjust = 0.5,
                                     vjust = 0.5), 
                   xmin = 0, xmax = 25, ymin = 0, ymax = 50) 

bird_plot = bird + bird_text + plot_layout(design = layout)

#################################
#          DOG PLOT             #
#################################

dog_png = readPNG("dog.png")
dog_color = "#DBCC83"

dog_text = ggplot() + 
  geom_textbox(aes(x =  30, y = 45, label = "<span style='font-size: 36pt;'><strong>Dogs</strong></span><br/><span style='font-size: 24pt;'>London Fire Brigade responded to <strong>1,194 incidents</strong> between 2009 and 2021. Dogs are often trapped when their head, leg, or paw become stuck in items near the household such as gates, fences, railways, or fox and rabbit holes.</span>"), fill = dog_color, box.colour = dog_color, width = unit(400, "pt"), hjust = 0, vjust = 1, family = "Bahnschrift") +
  scale_x_continuous(limits = c(0,100)) +
  scale_y_continuous(limits = c(0,50)) +
  theme_void() +
  theme(plot.background = element_rect(fill = dog_color, color = dog_color),
        panel.background = element_rect(fill = dog_color, color = dog_color)) +
  
  annotation_custom(grob = rasterGrob(dog_png, 
                                     interpolate=TRUE,
                                     hjust = 0.5,
                                     vjust = 0.5), 
                   xmin = 0, xmax = 25, ymin = 0, ymax = 50) +
  labs(caption = "Source: London.gov | @a_lipphardt") +
  theme(plot.caption = element_markdown(family = "Bahnschrift", 
                                    size = 18, 
                                    margin = margin(t = 5, b = 5, r = 20)))




dog_plot = dog + dog_text + plot_layout(design = layout)


#################################
#          FINAL PLOT           #
#################################

final_plot = cat_plot / bird_plot / dog_plot

```



# Saving Image(s)

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r save_plot, warning=FALSE}


ggsave(plot = cat_plot,
       filename = "cat_bigrams.png",
       width=15,
       height=5,
       type = "cairo",
       device = "png",
       dpi = 300)

ggsave(plot = bird_plot,
       filename = "bird_bigrams.png",
       width=15,
       height=5,
       type = "cairo",
       device = "png",
       dpi = 300)

ggsave(plot = dog_plot,
       filename = "dog_bigrams.png",
       width=15,
       height=5,
       type = "cairo",
       device = "png",
       dpi = 300)


ggsave(plot = final_plot,
       filename = "animal_rescues_bigrams.png",
       width=15,
       height=15,
       type = "cairo",
       device = "png",
       dpi = 300)

```

![There were 3,649 cat incidents between 2009 and 2021. Cats may often become trapped in areas of the house such as under the floor boards, in the washing machine, on window ledges, or in cars. There were 1,530 bird incidents between 2009 and 2021. Birds may often become trapped when their legs become stuck in netting or wire. Common locations include underneath railway bridges or rooftops. Incidents incldue both domesticated and wild birds. There were 1,194 dog incidents between 2009 and 2021. Dogs may often become trapped when their head, leg, or paw become stuck in items near the household such as gates, fences, railways, or fox and rabbit holes.](animal_rescues_bigrams.png)







