---
title: "Broadband Internet Access"
author: Developed by Anthony Lipphardt
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = 'README.md') })
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) event!

Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.

While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library("tidytuesdayR")
library("tidyverse")
library("glue")
library("ggtext")
library("ggalt")
library("ggthemes")


```

# Loading the Weekly Dataset

Download the weekly data and make available in the `broadband` object.

[Broadband](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-05-11/readme.md)

```{r Load}

# Loading Data for the First Time
# tuesdata <- tidytuesdayR::tt_load(2021, week = 20)
# broadband <- tuesdata$broadband
# write_csv(broadband, 'broadband.csv')

broadband <- read_csv('broadband.csv')

```


# Sample Data

Take an initial look at the format of the data available.

```{r Glimpse}

head(glimpse(broadband), 10)

```

# Wrangling

Explore the data and process it into prepared dataframes for visualization.

```{r Wrangle}

md_bb = broadband %>% 
  filter(ST == 'MD') %>%
  rename(COUNTY_ID = `COUNTY ID`,
         COUNTY_NAME = `COUNTY NAME`,
         AVAILABILITY = `BROADBAND AVAILABILITY PER FCC`,
         USAGE = `BROADBAND USAGE`) %>%
  mutate(COUNTY_NAME = str_remove(COUNTY_NAME, " County"),
         AVAILABILITY = as.numeric(AVAILABILITY) * 100,
         USAGE = as.numeric(USAGE) * 100,
         DISPARITY = AVAILABILITY - USAGE,
         COUNTY_NAME = fct_reorder(COUNTY_NAME, desc(DISPARITY)))
  

```


# Visualization(s)

Using your processed dataset, create your unique visualization(s).

```{r Visualize, results = FALSE, fig.width=16, fig.height=9}

color1 <- "#58AF4A"
color2 <- "#0171CE"

md_title = glue("<span style='color:{color1}; font-weight:bold;'>AVAILABILITY</span> vs. <span style='color:{color2}'>USAGE</span> of Broadband Internet in {md_bb[1,]$ST} Counties")

highest = md_bb %>% slice_max(DISPARITY, n=1)
lowest = md_bb %>% slice_min(DISPARITY, n=1)

plot1 = md_bb %>% ggplot() +
  
  geom_segment(aes(y = COUNTY_NAME, yend = COUNTY_NAME, x = 0, xend = 100),
               color = "#CDCDCD",
               size = 0.2) +
  
  geom_dumbbell(aes(y = COUNTY_NAME, x = AVAILABILITY, xend = USAGE),
                size = 2, 
                color = "#CDCDCD",
                size_x = 3,
                size_xend = 3,
                colour_x = color1,
                colour_xend = color2,
                show.legend = TRUE) +
  
  labs(title = md_title,
       x = "Percent of People within County",
       y = "") +
  
  scale_x_continuous(breaks=seq(0,100,10), limits=c(0,100)) + 
  scale_y_discrete(limits=rev) +

  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +  
    
  theme(  
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size=12, face = "bold", color = "#444444"),
    plot.title = element_markdown(size = 18, hjust = 0.5)
    ) +
  
  geom_label(aes(y = highest$COUNTY_NAME, x = highest$USAGE, label = "Highest Disparity"), nudge_x = -8, alpha = 0.1) +
  geom_label(aes(y = lowest$COUNTY_NAME, x = lowest$USAGE, label = "Lowest Disparity"), nudge_x = -8, alpha = 0.1)
  
plot1

  
```

# Saving Image(s)

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(plot = plot1,
  filename = "broadband-disparities-md-counties.png",
  width=16,
  height=9,
  device = "png")

```

