---
title: "Scooby Doo Episodes"
author: Developed by Anthony Lipphardt
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = 'README.md') })
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) event!

Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.

While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.

The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!

As such we encourage everyone of all skills to participate!

```{r setup, warning=FALSE, results = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidytuesdayR)
library(tidyverse)

library(lubridate)
library(scales)
library(glue)
library(ggtext)
library(Cairo)


library(extrafont)
loadfonts(device = "win", quiet = TRUE)

library(colormap)
library(viridis)
library(ggforce)
library(igraph)
library(ggraph)



```

# Loading the Weekly Dataset

Download the weekly data and make it available in the `scoobydoo` object.

[Scooby Doo Episodes](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-07-13/readme.md)

```{r Load}

# Loading Data for the First Time
tuesdata <- tidytuesdayR::tt_load(2021, week = 29)
scoobydoo <- tuesdata$scoobydoo
write_csv(scoobydoo, 'scoobydoo.csv')

scoobydoo <- read_csv('scoobydoo.csv') %>%
  mutate(monster_name = str_replace(monster_name, "Agent,4", "Agent 4,"),
         monster_name = str_replace(monster_name, "Agent,8", "Agent 8,"),
         monster_name = str_replace(monster_name, "Agent,9", "Agent 9,"), 
         monster_name = str_replace_all(monster_name, "Bogle", "Bogel"), 
         monster_name = str_replace_all(monster_name, "Demons", "Demon"),          
         monster_name = str_replace_all(monster_name, "Phantoms", "Phantom"),          
         monster_name = str_replace_all(monster_name, "Zombies", "Zombie"), 
         monster_name = str_replace_all(monster_name, "Frankenstein's Monster", "Frankenstein"),
         monster_name = str_replace_all(monster_name, "Evil Trees", "Evil Tree"), 
         monster_name = str_replace_all(monster_name, "Dr Trebla", "Dr. Trebla"),
         monster_name = str_replace_all(monster_name, "Trebal", "Trebla"), 
         monster_name = str_replace_all(monster_name, pattern = regex("Agent \\d+"), replacement = "Agent X"))

```


# Sample Data

Take an initial look at the format of the data available.

```{r Glimpse}

glimpse(scoobydoo)


```

# Wrangling and Exploration


Counts for several categoricals.

```{r categorical_counts}

categorical_count = function(tbl, column){
  tbl %>% count({{column}}, sort = TRUE)
}

categorical_count(scoobydoo, series_name)
categorical_count(scoobydoo, network)
categorical_count(scoobydoo, format)
categorical_count(scoobydoo, monster_gender)
categorical_count(scoobydoo, monster_real)
categorical_count(scoobydoo, setting_terrain)
categorical_count(scoobydoo, setting_country_state)
categorical_count(scoobydoo, non_suspect)
categorical_count(scoobydoo, arrested)
categorical_count(scoobydoo, culprit_gender)
categorical_count(scoobydoo, if_it_wasnt_for)
categorical_count(scoobydoo, and_that)
categorical_count(scoobydoo, door_gag)
categorical_count(scoobydoo, split_up)
categorical_count(scoobydoo, another_mystery)
categorical_count(scoobydoo, batman)


```

**monster_name** look's like it contains all monsters that appear in a given episode and from the looks of it there are some common monsters that have appeared in the series. Let's see whether we can unpack this column to get monsters that have appeared together.

NOTE: Plural versions of monster names have been resolved to singular (e.g. Demons to Demon, Zombies to Zombie, etc) and mispellings have been resolved. There are also listings for each of the individual Agent characters, which have been collapsed into a single Agent X for a more interesting graph.

```{r monster_name}


# Unpack monster_name
monster_names = scoobydoo %>% select(index, monster_name) %>% mutate(monster_name = str_split(monster_name, ",")) %>% unnest(monster_name) %>% distinct(index, monster_name)

# Number of episodes monster has appeared in
monster_name_counts = monster_names %>% filter(!(monster_name %in% c("NULL", ""))) %>% count(monster_name, sort=TRUE)

# Monsters that have appears more than once
monster_name_counts %>% filter(n > 1)

# Monsters that were paired together
monster_pairs = monster_names %>% inner_join(monster_names, by = "index") %>% 
  filter(monster_name.x != monster_name.y,
         monster_name.x < monster_name.y) %>% 
  unique()

# Top monster pairings
monster_pair_counts = monster_pairs %>% count(monster_name.x, monster_name.y, sort=TRUE)

# Unique monster names in pairings
unique_names = unique(c(monster_pairs$monster_name.x, monster_pairs$monster_name.y))

# Pull the top 50 monsters by number of appearances. 
unique_names_top = monster_name_counts %>% 
  filter(monster_name %in% unique_names) %>% 
  slice_max(n=50, n, with_ties = FALSE) %>%
  pull(monster_name)

# Now pull all pairings for those top 50 monsters
monster_pair_counts_top = monster_pair_counts %>% 
  filter( (monster_name.x %in% unique_names_top) & (monster_name.y %in% unique_names_top) ) %>%
  arrange(monster_name.x)


```



# Visualization

Create a chord diagram, with monster names arranged in a circle. Each point around the circle should relate to the number of episodes each monster appeared in. 

```{r chord}

# Scooby Doo Palette
scooby_blue = "#8FEAFC"
scooby_green = "#CAFF4D"
scooby_orange = "#FF4400"
scooby_brown = "#6F440F"


# Monster Info for Vertices - Sort in Alphabetical Order

monster_info = monster_name_counts %>% 
  filter(monster_name %in% unique_names_top) %>%
  mutate(strength = case_when(
    n < 3 ~ "Low",
    n < 10 ~ "Medium",
    TRUE ~ "High"),
    strength = as.factor(strength)) %>%
  arrange(desc(monster_name))


# Create graph based on pairings

monster_graph = monster_pair_counts_top %>% 
  graph_from_data_frame(directed = FALSE, 
                        vertices = monster_info)


# Number of vertices/monsters in graph
number_of_monsters = gorder(monster_graph)


# Calculate angle and hjust for each text label
angle = 360 * (seq(1, number_of_monsters) - 0.5) / number_of_monsters
angle_adj = ifelse(angle > 90 & angle<270, angle+180, angle)
hjust = ifelse(angle > 90 & angle<270, 1, 0)


# Create chord graph to show monster pairings and number of appearances
chord_graph = ggraph(monster_graph, layout = "circle") +
  
  # Green circle on inside of graph
  geom_circle(aes(x0 = 0, y0 = 0, r = 1), fill = scooby_green, color = "black", size = 1) +
  
  # Brown connections between monster pairs, more appearances result in thicker line
  geom_edge_link(aes(edge_width = log(n)), 
                 edge_colour=scooby_brown, 
                 edge_alpha = 1, 
                 show.legend = FALSE) +  
  
  # Orange circle/node for each monster
  geom_node_point(size = 8, fill = scooby_orange, colour = "black", pch = 21, 
                  show.legend = FALSE) +
  

  # Number of appearances for a given monster
  geom_node_text(aes(label=glue("{n}")), 
                 size=4, 
                 color="black", 
                 show.legend = FALSE,
                 family = "Segoe UI") +
  
  # Name of monster around outside of graph  
  geom_node_text(aes(label=glue("     {name}     "), 
                     angle = angle_adj, 
                     hjust = hjust), 
                 size=5, 
                 color="black", 
                 show.legend = FALSE,
                 family = "Rockwell") +
  
  # Connections/edges should fall within this range scaled by number of appearances
  scale_edge_width(range = c(1, 3)) +
  
  theme_void() +
  
  # Expand limits to add padding for node labels
  expand_limits(x = c(-1.5, 1.5), y = c(-1.5, 1.5)) +
  
  labs(title = "The Monsters of Scooby Doo",
       subtitle = "With more than 600 mysteries in the history of the franchise, the gang has frequently encountered the same monster on multiple occasions. The following graph shows the Top 50 monsters by number of appearances (indicated in the orange nodes) and whether they've appeared alongside other monsters (indicated by brown edges). Thicker edges in the graph represent more appearances for a given pair of monsters.",
       caption = "<strong>Source:</strong> ScoobyPedia | <strong>Visualization:</strong> @a_lipphardt") +
  
  theme(plot.background = element_rect(fill = scooby_blue, color = scooby_blue),
        
        plot.title = element_textbox_simple(family = "Cooper Black",
                                            halign = 0.5,
                                            size = 20,
                                            margin = margin(t=10,b=10)),
        
        plot.subtitle = element_textbox_simple(family = "Segoe UI",
                                            halign = 0.5,
                                            size = 16,
                                            margin = margin(b=20, 
                                                            l = 20, 
                                                            r = 20)),
        
        plot.caption = element_textbox_simple(family = "Segoe UI Semilight",
                                              size = 14,
                                              halign = 1,
                                              margin = margin(r = 10, t = 20, b = 5)))

```




# Saving Image(s)

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r save_plot, warning=FALSE}

ggsave(plot = chord_graph,
       filename = "scooby-doo.png",
       width=12,
       height=12,
       type = "cairo",
       device = "png",
       dpi = 300)

```

![The Monsters of Scooby Doo. With more than 600 mysteries in the history of the franchise, the gang has frequently encountered the same monster on multiple occasions. The following graph shows the Top 50 monsters by number of appearances (indicated in the orange nodes) and whether they've appeared alongside other monsters (indicated by brown edges). Thicker edges in the graph represent more appearances for a given pair of monsters.](scooby-doo.png)





